|

### 2019-08-01 article/article.Rmd
small changes

However, accountability systems can take many forms. One promising model for democratic oversight is _bottom-up monitoring_, in which citizens receive information about the shortcomings of a given project to help them evaluate and pressure underperforming public officials [@kosack2014does; @molina2016community; @raffler2018weakness]. Proponents suggest bottom-up accountability is effective because 1) constituents have first-hand information about the outcomes of local policies; 2) citizens have incentives to avoid corruption that directly affects them; 3) policy-makers are particularly sensitive to social punishment from their own communities [@serra2011combining, 570]. In a nutshell, bottom-up accountability offers a potential solution to the principal-agent dilemma in public service by aligning the interests of politicians and bureaucrats with those of the constituency they serve [@barro1973control; @raffler2018weakness, 2].

Here we assess the impact of a technology-driven solution that lowers the costs of evaluating public works and punishing political representatives in Brazil. Developed by Transparência Brasil[^transparencia], the _Tá de Pé_ (TDP) mobile phone application allows Brazilian citizens to know the location of public school construction sites, check their completion status, and anonymously request information from competent authorities. TDP users can also take pictures of delayed works and submit them to a group of trained engineers for examination. If the engineers classify the construction as delayed, TDP sends a message to the mayor's office asking for completion estimates and explanations about the construction status. TDP has been online since April 2017 in 1030 Brazilian municipalities and was the winner of the 2016 Google Social Impact grant with more than 200 thousand popular votes[^google].

@@

We used TDP to conduct **two/three?** experimental interventions. We tested whether the app has a positive impact on **school completion rates, school construction status and complaints to competent authorities**. Contrary to our expectations, the TDP app has delivered mixed results. **One/Two** of the findings are positive and in line with our theoretical predictions. From August 2017 to February 2019, TDP increased the likelihood of school completion by 6.8 percent (SE = X.X, t-test = XXX). Additionally, we find that pressure over the bureaucracy increased the investments in the schools by 157.5 percent (SE = X.XX, t-test = XXX). In contrast, our results show the app had a null impact on school constructions statistics. Brazilian politicians did not respond to the citizen's requests placed via the application. Therefore, our findings raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies. **Nota**: _Reescrever essa última frase e colocar mais explicações sobre possíveis mecanismos, etc_

@@

The _Tá de Pé_[^tdp] cell phone application is an initiative carried out by Transparência Brasil to foster bottom-up accountability in the Brazilian public sector. More specifically, the main goal of the TDP project is to improve responsiveness in government education expenditures. The TDP app incentivises citizens to provide up-to-date information about unfinished schools in their neighbourhoods, and the data will be assessed by a group of independent specialists. In the case the construction is behind schedule, TDP provides a writing platform whereby citizens can contact their representatives quickly and anonymously. The app then sends a notification to the mayor's office, which has 15 days to reply. If they do not respond to the request, a notification goes to the federal government, making it harder to the municipality to access federal funds in the future. The motivation for the intervention is that information delivery causes individuals to closely monitor public works. This, in turn, results in better social outcomes as public agents become more responsive to community demands.

@@

We built the app in the first three months of 2017 and tested it in May 2017. The first stable Android version was deployed on Google Play on 14 August 2017. A version for iOS appeared six months later. A massive Facebook campaign started in order to publicise the app. Facebook is the most widely used social media network in Brazil with around 72 million users in 2017 [@statista2019facebook]. The campaigns were concentrated in October 2017, investing around BRL 14.333,74 (US$ 4.479,29), and attracting 2,028 new users in October 2017 only.

@@

Transparência Brasil established a partnership with the Brazilian branch of _Engineers without Borders_ (EWB), a non-governmental organisation dedicated to promote international development[^ewb], to provide unbiased evaluations of school completion rates. Their assessment is based on photos and GPS location sent through the app. All data are collected anonymously and neither organisation are able to identify which user uploaded the images. Information on users' alerts are stored on their mobile devices, so citizens can track whether their request has been answered by public authorities.

>>>>

However, accountability systems can take many forms. One promising model for democratic oversight is _bottom-up monitoring_, in which citizens receive information about the shortcomings of a given project to help them evaluate and pressure underperforming public officials [@kosack2014does; @molina2016community; @raffler2018weakness]. Proponents argue bottom-up accountability is effective because 1) constituents have first-hand information about the outcomes of local policies; 2) citizens have incentives to avoid corruption that directly affects them; 3) policy-makers are particularly sensitive to social punishment from their own communities [@serra2011combining, 570]. In a nutshell, bottom-up accountability offers a potential solution to the principal-agent dilemma in public service by aligning the interests of politicians and bureaucrats with those of the constituency they serve [@barro1973control; @raffler2018weakness, 2].

Here we assess the impact of a technology-driven solution that lowers the costs of evaluating public works and punishing political representatives in Brazil. Developed by Transparência Brasil[^transparencia], the _Tá de Pé_ (TDP) mobile phone application allows Brazilian citizens to know the location of public school construction sites, check their completion status and anonymously request information from competent authorities. TDP users can also take pictures of delayed works and submit them to a group of trained engineers for examination. If the engineers classify the construction as delayed, TDP allows users to send a message to the mayor's office asking for completion estimates and explanations about the construction status. TDP has been online since April 2017 in 1030 Brazilian municipalities and was the winner of the 2016 Google Social Impact grant with more than 200,000 popular votes[^google].

@@

We use the TDP app to conduct **two/three?** experimental interventions. We test whether the app has a positive impact on **school completion rates, school construction status and complaints to competent authorities**. Contrary to our expectations, our experiments have delivered mixed results. **One/Two** of the findings are positive and in line with our theoretical predictions. From August 2017 to February 2019, TDP increased the likelihood of school completion by 6.8 percent (SE = X.X, t-test = XXX). Additionally, we find that pressure over the bureaucracy increased the investments in the schools by 157.5 percent (SE = X.XX, t-test = XXX). In contrast, our results show the app has a null impact on school constructions statistics. Brazilian politicians did not respond to the citizen's requests placed via the application. Therefore, our findings raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies. **Nota**: _Reescrever essa última frase e colocar mais explicações sobre possíveis mecanismos, etc_

@@

The _Tá de Pé_[^tdp] cell phone application is an initiative carried out by Transparência Brasil to foster bottom-up accountability in the Brazilian public sector. More specifically, the main goal of the TDP project is to improve responsiveness in government education expenditures. The TDP app incentivises citizens to provide up-to-date information about unfinished schools in their neighbourhoods, and that information will be assessed by a group of independent specialists. In the case the construction is behind schedule, TDP provides a writing platform whereby citizens can contact their representatives quickly and anonymously. The app then prompts citizen to write a notification to the mayor's office, which has 15 days to reply. If they do not respond to the request, the app forwards the notification to the Brazilian federal government, making it harder to the municipality to access federal funds in the future. The motivation for the intervention is that information delivery causes individuals to closely monitor public works. This, in turn, results in better social outcomes as public agents become more responsive to community demands.

@@

We built the app in the first three months of 2017 and tested it in May 2017. The first stable Android version was deployed on Google Play on 14 August 2017. A version for iOS appeared six months later. A massive Facebook campaign started in order to publicise the app. Facebook is the most widely used social media network in Brazil with around 72 million users in 2017 [@statista2019facebook]. The campaigns were concentrated in October 2017 and cost US$ 4,479.29. Facebook advertisements attracted 2,028 new users to the platform in October 2017 only.

@@

Transparência Brasil established a partnership with the Brazilian branch of _Engineers without Borders_ (EWB), a non-governmental organisation dedicated to promote international development[^ewb], to provide unbiased evaluations of school completion rates. Their assessment is based on photos and GPS location sent through the app. All data are collected anonymously. Information on users' alerts are stored on their mobile devices, so citizens can track whether their request has been answered by public authorities.
